%% $Id$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[english]{article}
\usepackage[latin1]{inputenc}
\usepackage{babel}
\usepackage{verbatim}

%% do we have the `hyperref package?
\IfFileExists{hyperref.sty}{
   \usepackage[bookmarksopen,bookmarksnumbered]{hyperref}
}{}

%% do we have the `fancyhdr' or `fancyheadings' package?
\IfFileExists{fancyhdr.sty}{
\usepackage[fancyhdr]{latex2man}
}{
\IfFileExists{fancyheadings.sty}{
\usepackage[fancy]{latex2man}
}{
\usepackage[nofancy]{latex2man}
\message{no fancyhdr or fancyheadings package present, discard it}
}}

%% do we have the `rcsinfo' package?
\IfFileExists{rcsinfo.sty}{
\usepackage[nofancy]{rcsinfo}
\rcsInfo $Id$
\setDate{\rcsInfoLongDate}
}{
\setDate{ 2012/09/21}
\message{package rcsinfo not present, discard it}
}

\setVersionWord{Version:}  %%% that's the default, no need to set it.
\setVersion{=PACKAGE_VERSION=}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{Name}{1}{hpcviewer}{The HPCToolkit Performance Tools}{The HPCToolkit Performance Tools}{hpcviewer:\\ Interactive Presentation of Performance}

The Java-based \Prog{hpcviewer} interactively presents program performance in a top-down fashion.

\end{Name}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synopsis}

Command-line usage:\\
\SP\SP\SP\Prog{hpcviewer} \oOpt{options} \oOpt{hpctoolkit-database}

GUI usage:\\
\SP\SP\SP Launch \File{hpcviewer} and open the Experiment database \oOpt{hpctoolkit-database}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Description}

The Java-based \Prog{hpcviewer} interactively presents program-performance Experiment databases in a top-down fashion.
Since Experiment databases are self-contained, they may be relocated from a cluster for visualization on a laptop or workstation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Arguments}

\begin{Description}
\item[\Arg{hpctoolkit-database}] An HPCToolkit Experiment database, which is the result of executing \Prog{hpcprof}, \Prog{hpcprof-mpi} or \Prog{hpcprof-flat}.
\end{Description}

%Default values for an option's optional arguments are shown in \{\}.

\subsection{Options}

\begin{Description}

\item[\Opt{-consolelog}]
Send log entries to a console in addition to a log file.

\item[\Opt{-debug}]
Log additional information about plug-in dependency problems.

\end{Description}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Detailed Description}

\subsection{Profile View}

\Prog{hpcviewer} supports three principal views of an application's performance data.  Both inclusive (costs of a procedure including all its callees) and exclusive costs (costs excluding callees) are reported.

\begin{itemize}
  \item \textbf{Calling context view.}
A top-down view that represents the dynamic calling contexts (call paths) in which costs were incurred.
Using this view, one can explore performance measurements of an application in a top-down fashion to understand the costs incurred by calls to a procedure in a particular calling context.

  \item \textbf{Callers view.}
This bottom-up view enables one to look upward along call paths.
The view apportions a procedure's costs to its caller and, more generally, its calling contexts.
This view is particularly useful for understanding the performance of software components or procedures that are used in more than one context.

  \item \textbf{Flat view.}
This view organizes performance measurement data according to the static structure of an application.
All costs incurred in \emph{any} calling context by a procedure are aggregated together in the flat view.
\end{itemize}


\subsubsection{Panes}

The browser window is divided into three panes:

\begin{itemize}
  \item \textbf{Source pane.} The source code associated with the current entity selected in the navigation pane.
Selecting any entity in the navigation pane will cause the source pane to load the corresponding file, scroll to and highlight the line corresponding to the selection.

\item \textbf{Navigation pane.}
The navigation pane presents a hierarchical tree-based structure that is used to organize the presentation of an application's performance data.
Entities that occur in the navigation pane's tree include load modules, files, procedures, procedure activations, inlined code, loops, and source lines.
Entities can be selected; and children can be revealed or concealed.

The nature of the entities in the navigation pane's tree structure depends upon whether one is exploring the calling context view, the callers view, or the flat view of the performance data. 

\begin{itemize}
\item In the calling context view, entities in the navigation tree represent procedure activations, inlined code, loops, and source lines. 
While most entities link to a single location in source code, procedure activations link to two: the call site from which a procedure was called and the procedure itself.
\item In the callers view, entities in the navigation tree are procedure activations. Unlike procedure activations in the calling context tree view in which call sites are paired with the called procedure, in the caller's view, call sites are paired with the calling procedure to facilitate attribution of costs for a called procedure to multiple different call sites and callers.
\item In the flat view, entities in the navigation tree correspond to source files, procedure call sites (which are rendered the same way as procedure activations), loops, and source lines.
\end{itemize}


The header above the navigation pane contains flatten and zoom control.
\begin{itemize}
\item \textbf{Up arrow.} Zooms to show only information for the selected line and its descendants.
\item \textbf{Down arrow.} Zooms out (reversing a prior zoom operation).
\item \textbf{Hot path}. The button activates "\emph{hot path analysis}", which automatically expands the scopes along the hot path for the selected metric in the subtree rooted at the selected scope. It is a convenient way to find a performance bottleneck of a given metric.
\item \textbf{Derived metric}. Creation of a new metric based on spreadsheet-like mathematical formulae.
\item \textbf{Filter metrics}. Show or hide metrics.
\item \textbf{CSV export}. Export information from the current view to a comma separated values (CSV) file.
\item \textbf{Bigger font}. Change the current font with a bigger size font.
\item \textbf{Smaller font}. Change the current font with a smaller size font.
\item \textbf{Flatten.} (flat view only) Flattening (the icon that shows a tree node with a slash through it) replaces each top-level scope shown with its children. If a scope has no children (i.e., it is a leaf), the node will remain in the view. This flattening operation is useful for relaxing the strict hierarchical view so that peers at the same level in the tree can be viewed and ranked together. For instance, this can be used to hide procedures in the flat view so that outermost loops can be ranked and compared to one another.
\item \textbf{Unflatten.} (flat view only) The inverse of the flatten operation, making an elided node visible again.

\end{itemize}

\item  \textbf{Metric pane.}
The metric pane displays one or more performance metrics associated with entities to the left in the navigation pane.
Entities in the tree view of the navigation pane are sorted at each level of the hierarchy by the metric in the selected column.

\end{itemize}


\subsubsection{Thread-Centric Plots}

\Prog{hpcviewer} can display plot graphs of thread-level metric
values.  This is useful for quickly assessing load imbalance across
processes and threads.

To create a graph, use the calling context view and select a scope by
left-clicking a line in the navigation pane.
Then, right-click the selected scope to show the context menu.  (This
menu begins with `Zoom-in' and `Zoom-out.')
At the bottom of the context menu is a list of metrics that
\Prog{hpcviewer} can graph.
Each metric contains a sub-menu that lists the three different types
of graphs that \Prog{hpcviewer} can plot.
The \emph{Plot} graph sorts the processes by process and thread
number.
The \emph{Sorted plot} graph sorts the processes by metric value.
And the \emph{Histogram} graph shows a bar graph of the
frequency of metric values.

Note: currently, these plot graphs are available only with databases
created by \Prog{hpcprof-mpi} and not by \Prog{hpcprof}.
See the section on ``\emph{Plotting Graphs of Thread-level Metric Values}''
in the User's Manual for more description and a view of a plot graph.



\subsection{Trace view}

The view interactively presents program traces in a top-down fashion. 
It comprises of three different parts. 


\begin{itemize}
\item Main view (left, top):
  This is hpctraceviewer's primary view.
  This view, which is similar to a conventional process/time (or space/time) view, shows time on the horizontal axis and process (or thread) rank on the vertical axis; time moves from left to right.
  Compared to typical process/time views, there is one key difference.
  To show call path hierarchy, the view is actually a user-controllable slice of the process/time/call-path space.
  Given a call path depth, the view shows the color of the currently active procedure at a given time and process rank.
  (If the requested depth is deeper than a particular call path, then hpctraceviewer simply displays the deepest procedure frame and, space permitting, overlays an annotation indicating the fact that this frame represents a shallower depth.)
  hpctraceviewer assigns colors to procedures based on (static) source code procedures.
  Although the color assignment is currently random, it is consistent across the different views.
  Thus, the same color within the Trace and Depth Views refers to the same procedure.
  The Trace View has a white crosshair that represents a selected point in time and process space.
  For this selected point, the Call Path View shows the corresponding call path.
  The Depth View shows the selected process.

\item Depth view (left, bottom):
  This is a call-path/time view for the process rank selected by the Trace view's crosshair.
  Given a process rank, the view shows for each virtual time along the horizontal axis a stylized call path along the vertical axis, where `main' is at the top and leaves (samples) are at the bottom.
  In other words, this view shows for the whole time range, in qualitative fashion, what the Call Path View shows for a selected point.
  The horizontal time axis is exactly aligned with the Trace View's time axis; and the colors are consistent across both views.
  This view has its own crosshair that corresponds to the currently selected time and call path depth.

\item Summary view (same location as depth view on the left-bottom part):
  The view shows for the whole time range dislayed, the proportion of each subroutine in a certain time.
  Similar to Depth view, the time range in Summary reflects to the time range in the Trace view. 

\item Call view (right, top):
  This view shows two things: (1) the current call path depth that defines the hierarchical slice shown in the Trace View; and (2) the actual call path for the point selected by the Trace View's crosshair.
  (To easily coordinate the call path depth value with the call path, the Call Path View currently suppresses details such as loop structure and call sites; we may use indentation or other techniques to display this in the future.)

\item Mini map view (right, bottom):
  The Mini Map shows, relative to the process/time dimensions, the portion of the execution shown by the Trace View.
  The Mini Map enables one to zoom and to move from one close-up to another quickly.

\end{itemize}


% ===========================================================================

\subsubsection{Main view}

Main view is divided into two parts: the top part which contains \emph{action pane} and the \emph{information pane}, and the main view which displays the traces. 

The buttons in the action pane are the following:
\begin{itemize}

\item Home : Resetting the view configuration into the original view, i.e., viewing traces for all times and processes.
\item Horiontal zoom in / out : Zooming in/out the time dimension of the traces. 
\item Vertical zoom in / out : Zooming in/out the process dimension of the traces.
\item Navigation buttons : Navigating the trace view to the left, right, up and bottom, respectively. It is also possible to navigate with the arrow keys in the keyboard. Since Trace view does not support scrool bars, the only way to navigate is through navigation buttons (or arrow keys).
\item Undo : Canceling the action of zoom or navigation and returning back to the previous view configuration.
\item Redo : Redoing of previously undo change of view configuration.
\item save  / a view configuration : Saving/loading a saved view configuration. 
A view configuration file contains the information of the current dimension of time and process, the depth and the position of the crosshair. 
It is recommended to store the view configuration file in the same directory as the database to ensure that the view configuration file matches well with the database since the file does not store which database it is associated with. 
Although it is possible to open a view configuration file which is associated from different database, it is highly not recommended since each database has different time/process dimensions and depth.


\end{itemize}

The information pane contains some information concerning the range status of the current displayed data.
\begin{itemize}
 \item Time Range. The information of current time-range (horizontal) dimension. 
 \item Process Range. The information of current process-range (vertical) dimension. 
 \item Cross Hair. The information of current crosshair position in time and process dimensions. 
\end{itemize}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Arguments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Examples}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Notes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{See Also}

\HTMLhref{hpctoolkit.html}{\Cmd{hpctoolkit}{1}}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Version}

Version: \Version\ of \Date.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{License and Copyright}

\begin{description}
\item[Copyright] \copyright\ 2002-2020, Rice University.
\item[License] See \File{README.License}.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors}

\noindent
Laksono Adhianto \\
John Mellor-Crummey \\
Nathan Tallent \\
Rice HPCToolkit Research Group \\
Email: \Email{hpctoolkit-forum =at= rice.edu} \\
WWW: \URL{http://hpctoolkit.org}.

\LatexManEnd

\end{document}

%% Local Variables:
%% eval: (add-hook 'write-file-hooks 'time-stamp)
%% time-stamp-start: "setDate{ "
%% time-stamp-format: "%:y/%02m/%02d"
%% time-stamp-end: "}\n"
%% time-stamp-line-limit: 50
%% End:

